FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2017-10-19 NewVersion/sections/Overview.tex
small fix
Reductions

// main program
new DrawableSafeDeck().shuffleAndDraw()
\end{lstlisting}\vspace{3pt}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck|
is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver
is replaced by the object \lstinline|new DrawableSafeDeck()|.

From the perspective of \lstinline|DrawableSafeDeck|, the \lstinline|draw| method seems to be ambiguous since \lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from both \lstinline|SafeDeck| and \lstinline|Drawable|. But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch. Intuitively, it 
seems like we need to choose dynamic dispatch but also make use of static type
information, since the type of this-reference (\lstinline|Deck|)
specifies the branch to use unambiguously.

@@

\noindent We show the rough reductions here (without our techniques):
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [new DrawableSafeDeck().shuffle(), new DrawableSafeDeck().draw()]
-> [..., ambiguous call!}]
\end{lstlisting}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck| is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver 
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the \lstinline|draw| method is ambiguous since \lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from both \lstinline|SafeDeck| and \lstinline|Drawable|. But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch. Here is the rough reductions in our model that accept this call: 
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [..., (Deck)new DrawableSafeDeck().draw()]
-> [..., SafeDeck.draw()]
\end{lstlisting}
We make use of the branch information by adding the cast \lstinline|Deck|, since the type of this-reference (\lstinline|Deck|) specifies the branch to use unambiguously. 
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. This is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned. The details of how this is done will be explained in Section~\ref{sec:formalization}.

>>>>

// main program
new DrawableSafeDeck().shuffleAndDraw()
\end{lstlisting}\vspace{3pt}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck|
is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the
\lstinline|draw| method seems to be ambiguous since
\lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from
both \lstinline|SafeDeck| and \lstinline|Drawable|. 
Indeed, using the a reduction model similar to FJ~\cite{} (where static types tracked), the 
reduction steps would roughly be:
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [new DrawableSafeDeck().shuffle(), new DrawableSafeDeck().draw()]
-> [..., ambiguous call!}]
\end{lstlisting}
When the \lstinline|DrawableSafeDeck| object calls \lstinline|shuffleAndDraw|, the implementation in \lstinline|Deck| is dispatched. But then \lstinline|shuffleAndDraw| invokes ``\lstinline|this.draw()|'', and at this point, the receiver 
is replaced by the object \lstinline|new DrawableSafeDeck()|.
From the perspective of \lstinline|DrawableSafeDeck|, the \lstinline|draw| method is ambiguous since \lstinline|DrawableSafeDeck| inherits two \lstinline|draw| method from both \lstinline|SafeDeck| and \lstinline|Drawable|. But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch.

\begin{comment}
But ideally we would
like \lstinline|shuffleAndDraw| to invoke \lstinline|SafeDeck.draw|
because they belong to the same class hierarchy branch. Intuitively, it 
seems like we need to choose dynamic dispatch but also make use of static type
information, since the type of this-reference (\lstinline|Deck|)
specifies the branch to use unambiguously.
\end{comment}

@@

Here are the reductions performed in \MIM{} to accept this call: 
\begin{lstlisting}
new DrawableSafeDeck().shuffleAndDraw() 
-> [..., (Deck)new DrawableSafeDeck().draw()]
-> [..., SafeDeck.draw()]
\end{lstlisting}
Notably, we make use of the branch information by adding the cast \lstinline|Deck|, since the type of this-reference (\lstinline|Deck|) specifies the branch to use unambiguously. 
During the runtime, when \lstinline|this.draw()| is called,
the dynamic type of the receiver is \lstinline|DrawableSafeDeck|, but the static type is actually
\lstinline|Deck|. This is an instance of \textit{hierarchical invocation}, which can be read as 
``finding the most specific \lstinline|draw| method along path \lstinline|Deck|''. The meaning of ``along path
\lstinline|Deck|'' not only implies what \lstinline|Deck| inherits, but also the domain of its subtypes up to the
dynamic type \lstinline|DrawableSafeDeck|. Finally \lstinline|SafeDeck.draw| is correctly returned. The details of how this is done will be explained in Section~\ref{sec:formalization}.
